Validating Cell Tower Clustering for Mobility Prediction

### 2012-12-12 docs/paper589.tex
Stationary oscillation example.

%\documentclass[a4paper,10pt]{article}
\documentclass[letterpaper, 11pt, conference]{ieeeconf}
\usepackage{amsmath} % For 'cases' curly brace
\usepackage{amsfonts} % For bold number set font
\usepackage{color}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{amssymb}
\usepackage{url}

@@

\section{Introduction}
\label{sec:intro}
\xxx{Talk about mobility prediction in mobile networks. Discuss some of the potential benefits -- some of them were discussed in the
ATT leap graph paper. Maybe do a little research on this.}

In this project, we address one of the challenges faced in mobility prediction using GSM cell tower connection logs. In our context, mobility prediction in cellular
towers can be defined as the ability to predict a user's physical movement by examining the history of cell tower connections. This is very important
for mobile network operators, as several network optimizations are possible if they can forecast with reasonable accuracy
where a user will be in the future \cite{LeapGraph} \cite{Liu97anoptimal}. By being able to pre-allocate resources to towers that are most likely to
be connected to by the user, network operators can improve latency and uptime. This is increasingly important as the amount of data in cellular networks
increases: indeed, in 2010 the amount of data traffic surpassed voice traffic for the first time, and it is expected to double annually in the near future
\cite{EricssonData}.

@@

The heart of every mobility prediction system based on cell tower connections is to infer movement based on which cell the user's phone is connected to. The goal
is to predict which tower the user is more likely to connect to next. A major challenge in this is dealing with the \textit{oscillation effect}, which happens when a \textit{stationary} user's phone switches between two or more towers. This is a natural consequence of the radio medium and the fact that towers very often have overlapping ranges. In order to accurately predict a user's mobility, these outliers have to be removed so that fluctuations can be separated from actual user movement.

@@

@@

\xxx{Reproduce the stationary phone example in the leap graph paper.}

@@

There are two main papers that tackle the problem of oscillation effects: the Mobility Profiler paper, by Bayir et al. \cite{mobilityprofiler}, and the Leap Graph paper by Duffield et al \cite{LeapGraph}. Ideally we would want to evaluate both of them, but the leap graph algorithm, which was published by researchers at AT\&T, makes use of the active set of connections. The active set contains all towers that are within range of the phone's antenna at any given time. Unfortunately, active set logs are not public which makes it impossible for us to evaluate the leap graph algorithm.

>>>>

%\documentclass[a4paper,10pt]{article}
\documentclass[letterpaper, 11pt, conference]{ieeeconf}
\usepackage{amsmath} % For 'cases' curly brace
\usepackage{amsfonts} % For bold number set font
\usepackage{color}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{amssymb}
\usepackage{url}
\usepackage{graphicx}
\usepackage[font=bf]{caption}

@@

\section{Introduction}
\label{sec:intro}

In this project, we address one of the challenges faced in mobility prediction using GSM cell tower connection logs. In our context, mobility prediction in cellular
towers can be defined as the ability to predict a user's physical movement by examning the history of cell tower connections. This is very important
for mobile network operators, as several network optimizations are possible if they can forecast with reasonable accuracy
where a user will be in the future \cite{LeapGraph} \cite{Liu97anoptimal}. By being able to pre-allocate resources to towers that are most likely to
be connected to by the user, network operators can improve latency and uptime. This is increasingly important as the amount of data in cellular networks
increases: indeed, in 2010 the amount of data traffic surpassed voice traffic for the first time, and it is expected to double annually in the near future
\cite{EricssonData}.

@@

The heart of every mobility prediction system based on cell tower connections is to infer movement based on which cell the user's phone is connected to. The goal
is to predict which tower the user is more likely to connect to next. A major challenge in this is dealing with the \textit{oscillation effect}, which happens when a \textit{stationary} user's phone switches between two or more towers. Figure~\ref{fig:stationary} shows an example of a stationary phone that connects to different cell towers. This is a natural consequence of the radio medium and the fact that towers very often have overlapping ranges. In order to accurately predict a user's mobility, these outliers have to be removed so that fluctuations can be separated from actual user movement.

@@

\begin{figure}
\centering
\includegraphics[width=0.15\textwidth]{figs/stationary_oscillation}
\caption{An example of actual cell tower connection logs for a stationary phone, slightly modified from~\cite{LeapGraph}. The time of the sample is on the left, and an identifier for the cell tower is on the right. We can see that the phone connects to several towers, even though it does not move. }
\label{fig:stationary}
\end{figure}

@@

@@

There are two main papers that tackle the problem of oscillation effects: the Mobility Profiler paper, by Bayir et al. \cite{mobilityprofiler}, and the Leap Graph paper by Duffield et al \cite{LeapGraph}. Ideally we would want to evaluate both of them, but the leap graph algorithm, which was published by resarchers at AT\&T, makes use of the active set of connections. The active set contains all towers that are within range of the phone's antenna at any given time. Unfortunately, active set logs are not public which makes it impossible for us to evaluate the leap graph algorithm.
